---
knit: "bookdown::render_book"
---

# (APPENDIX) Appendix {-} 

# Appendix

## Bayes Factor Robustness Checks {#bfrc}

**Hypothesis 1 --** The checkmarks in Table \@ref(tab:apx-bfrc-h1) depict differences in prior specifications over the priors chosen in the reported models in the results section. Changes to _df = 7_ relate to all model parameters, _Slope = 0_ relates to the location/mean of main effects and truncation means a lower and upper bound of -10 and 10 for the prior on the intercept. 

(ref:apx-bfrc-h1) Bayes Factor robustness check for H1: the same model (including both effects without interaction) run against the intercept only model with different priors.

```{r apx-bfrc-h1}
options(knitr.kable.NA = "")
readr::read_delim("models/h1-bfrc.csv") %>% 
  kbl(escape = FALSE, booktabs = TRUE, align = "c",
      caption="(ref:apx-bfrc-h1)") %>% 
  kable_styling(latex_options = "striped", full_width = FALSE,
                font_size = 10) %>%
  row_spec(7, hline_after = TRUE) %>%
  row_spec(8, bold = TRUE)
```

**Hypothesis 2 --** The same robustness check has been done for the model to analyse the training phase, displayed in table \@ref(tab:apx-bfrc-h2). Changes in prior assumptions are: changing the degrees of freedom of the student distributed parameters to 7 (same as before), assuming an intercept of 0 (that is, participants are just guessing) and assuming a slope of 0.4 (overall increase in learning accuracy by 10% by using rule instructions).

(ref:apx-bfrc-h2) Bayes Factor robustness check for H2.

```{r apx-bfrc-h2}
# options(knitr.kable.NA = "")
readr::read_delim("models/h2-bfrc.csv") %>% 
  kbl(escape = FALSE, booktabs = TRUE, align = "c",
      caption="(ref:apx-bfrc-h2)") %>% 
  kable_styling(latex_options = "striped", full_width = FALSE,
                font_size = 10) %>%
  row_spec(7, hline_after = TRUE) %>%
  row_spec(8, bold = TRUE)
```


## Individual Responses

Figure \@ref(fig:apx-transfer) shows each participants response to every stimulus in the transfer phase. Figure \@ref(fig:apx-probability) shows each participants certainty of a stimulus belonging to category Grot in the second run of the transfer phase.

```{r apx-transfer, fig.cap="Individual responses in transfer phase", out.width="100%", out.height="80%", cache=TRUE}
transfer %>% 
  ggplot(aes(img_x, img_y, fill = response)) +
    facet_wrap(~subj_id) +
    geom_tile(size = .2, color = "white") +
    scale_fill_manual(values = c("#ac0634", "#525252")) +
    theme_subjects +
    theme(legend.position = "none",
          strip.text = element_blank(),
          panel.spacing = unit(2, "pt"))
```

```{r apx-probability, fig.cap="Individual probability judgements of any given stimulus belonging to category Grot", out.width="100%", out.height="80%", cache=TRUE}
transp %>% 
  ggplot(aes(img_x, img_y, fill = prob)) +
    facet_wrap(~subj_id) +
    geom_tile(size = .2, color = "white") +
    scale_fill_viridis_c(
      # guide = guide_colorbar(
      #   title.position = "top",
      #   title.hjust	= .5,
      #   label.position = "bottom"
      # )
    ) +
    theme_transfer +
    theme(legend.position = "none",
          legend.key.height = unit(.3, "lines"),
          legend.text = element_text(),
          strip.text = element_blank(), # element_text(size = 8),
          panel.spacing = unit(2, "pt"))
```

